---
title: "T1D_MS"
author: "Sandeep Singh"
date: "Last Updated: `r format(Sys.time(), '%d, %B, %Y at %H:%M')`"
output: md_document
always_allow_html: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r}
# This document depends on the following packages:
library(rlang)
library(dplyr)
library(tibble)
library(readxl)
library(DT)
library(tidyr)
library(data.table)
library(ggplot2)
library(ggrepel)
library(forcats)
library(scales)
library(grid)
library(purrr) # to append multiple dataframes
library(plotly)
library(VennDiagram)
library(RColorBrewer)
library(png)
```

* raw_data.txt file was generated from previous researches on Type-1-diabetes and multiple sclerosis.  
* This file contains only 119 SNPs as common markers for T1D and MS. 

## File preparation
```{r}
T1D_MS_original_119 <- read.table("C:/SANDEEP/MAIN/CAREER/My_papers/PAPERS_in_PROGRESS/T1D_MS/T1D_MS_new/T1D.txt",
                                  header = F) 

colnames(T1D_MS_original_119) <- c("SNP")
# used to proivde column name

head(T1D_MS_original_119)
```

* A total of 119 SNPs.
* This file was used to create ANNOVAR formatted file.

### File preparation by ANNOVAR
* Command used in ANNOVAR: 

convert2annovar.pl -format rsid T1D.txt -dbsnpfile humandb/hg19_snp138.txt > T1D_MS_formatted

* 119 SNPs records input; 110 SNPs records as output, 9 SNPs missing

```{r}
# duplicate SNPs due to HLA region were removed from ANNOVAR file
T1D_MS_original_110 <- read.table("file:///C:/SANDEEP/MAIN/CAREER/My_papers/PAPERS_in_PROGRESS/T1D_MS/T1D_MS_new/T1D_MS_formatted.txt",
                                  header = F)

datatable(T1D_MS_original_110, class = 'cell-border stripe', 
          caption = 'Table 2: Annovar formatted file.',
          options = list(pageLength = 5, autowidth = TRUE,
  initComplete = JS(
    "function(settings, json) {",
    "$(this.api().table().header()).css({'background-color': '#000', 'color': '#4c4c4c'});",
    "}")
))
```
